Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-12 paper/comoving-followup.tex
crushed the intro re-write. fixes #5

Widely-separated, coeval pairs or multiplets of stars --- i.e. wide binaries,
multiplets, or former binaries or multiplets --- are important tracers of many
processes that formed and continue to shape the Milky Way.
For example, they have been used to place limits on the mass distribution and
population of massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to
study the dynamical outcomes of star formation processes (e.g.,
\citealt{Reipurth:2012}), and to calibrate stellar models and inferences across
the main sequence (e.g., \citealt{Brewer:2016}).
These examples represent a small subset of a much broader range of applications
(see, e.g., \citealt{Chaname:2007}).

At smaller separations ($\lesssim 0.1~\pc$), comoving pairs of stars are likely
bound wide binaries and are therefore coeval.
At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
tenuously bound stars, recently disrupted binaries, moving groups or dynamical
features, and the final products of dissolving stellar associations or open
clusters.
However, even at separations of $\sim 0.1~\pc$, we typically lack the precision
to determine the binding energy of a system and therefore can't determine its
true dynamical state; because of this ambiguity, we refer to both bound and
unbound systems collectively as ``comoving stars.''

Comoving stars are identified by searching for approximately cospatial stars
that appear to be comoving either in full-space or in sky-projected velocities.
Primarily thanks to large astrometric surveys and catalogs
(\citealt{ESA:1997,Lepine:2005,Gaia-Collaboration:2016}) and radial-velocity
surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving pairs
(\citealt{Shaya:2011}) and over 10,000 candidate systems
(\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,Oelkers:2017,
Andrews:2017}) have already been discovered.
Radial velocity follow-up is still needed to confirm the majority of these
pairs, though \gaia\ \DR{2} will provide radial velocities for the brightest
stars ($G < 12$; \citealt{Gaia-Collaboration:2016}).

It has long been recognized that the observed period or separation distribution
of binary stars encapsulates information about star formation and dynamical
processes of the Galaxy (\citealt{Opik:1924}).
[If the initial distribution power law, break and steepens from perturbations]
[a bit of theory about wide binaries...predictions for wide binary separations?]
(Weinberg87)

Using candidate comoving pairs selected based on their proper motions alone,
several groups have measured the separation distribution of comoving stars at
angular separations $\Delta \theta \lesssim 1^\circ$ where the false-positive
rate is expected to be lower (\citealt{Chaname:2004,Lepine:2007,Sesar:2008});
these are likely predominantly wide binaries.
In most cases, the angular or projected separation distribution is fit using a
decreasing power-law, and the power-law index, $\alpha$, is found to be $\alpha
\approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
0.65~\pc$ at a distance of $150~\pc$).
Tentative reports of a steepening or cutoff of the separation distribution at
larger separations have been claimed (e.g., \citealt{Yoo:2004,Quinn:2009}), but
these claims are limited by small numbers of confirmed pairs at large
separations.

While the separation distribution of wide binaries appears to and is expected to
fall off rapidly with increasing separation, predictions for the separation
distribution of \emph{comoving stars} are less clear.
Early work suggested ionized pairs don't matter much...
[But including tidal field...] (jiang and tremaine)
[But still haven't included perturbers...]
[Tentative evidence of this population in (shaya and olling)]
[Need larger samples of confirmed large-separation pairs]

In this work, we present a catalog of confirmed comoving star pairs discovered
through radial velocity follow-up of candidate comoving stars selected from
inferences using their astrometry alone (\citealt{Oh:2017}).
This work serves as a pilot study for data expected in upcoming data releases
from the \gaia\ mission.
We describe the data, observations, and spectroscopic reduction in
\sectionname~\ref{sec:data}.
If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
and \sectionname~\ref{sec:discussion} where we present and discuss our main
results.

>>>>

Widely-separated, coeval pairs or multiplets of stars --- i.e. wide binaries,
multiplets, or former binaries or multiplets --- are important tracers of many
processes that formed and continue to shape the Milky Way.
They have been used to place limits on the mass distribution and population of
massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to study the
dynamical outcomes of star formation processes (e.g., \citealt{Reipurth:2012}),
and to calibrate stellar models and inferences across the main sequence (e.g.,
\citealt{Brewer:2016}).
These examples represent a small subset of a much broader range of applications
(see, e.g., \citealt{Chaname:2007}).

At smaller separations ($\lesssim 0.1~\pc$), comoving pairs of stars are likely
bound wide binaries.
It has long been recognized that the observed period or separation distribution
of these low-binding-energy binaries encapsulates information about the
Galactic mass distribution (\citealt{Opik:1924,Oort:1950,Bahcall:1985}).
In particular, early simulations of populations of wide binaries showed that
the expected separation distribution, $p(s)$, for $0.01 < s < 1~\pc$ tends to
follow a power law, $p(s) \propto s^{-\alpha}$ for $\alpha \approx 1$--$2$,
with a possible steep break that depends on the details of the population of
massive perturbers in the Galaxy, i.e. giant molecular clouds (GMCs), black holes, etc. (\citealt{Weinberg:1987}).

At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
tenuously bound stars, recently disrupted binaries, moving groups or resonant
features, and the final products of dissolving stellar associations or
clusters.
However, even at separations of $\sim 0.1~\pc$, we typically lack the precision
to determine the binding energy of a system and therefore can't determine its
true dynamical state; because of this ambiguity, we refer to both bound and
unbound widely-separated ($\gtrsim 10^{-3}~\pc$) systems collectively as
``comoving'' systems.

Primarily thanks to large astrometric surveys and catalogs
(\citealt{ESA:1997,Lepine:2005,Gaia-Collaboration:2016}) and radial-velocity
surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving star
pairs (\citealt{Shaya:2011}) and over 10,000 candidate systems
(\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,
Oelkers:2017, Andrews:2017}) have been discovered over the last several decades (see also \citealt{Chaname:2007} and references therein).
These pairs are identified by searching for approximately cospatial stars that
appear to be comoving either in sky-projected velocities --- we refer to these
as ``comoving pair candidates'' --- or using full-space velocity information --- we refer to these as ``comoving pairs.''
Radial velocity follow-up is still needed to confirm the majority of the known
candidates, though \gaia\ \DR{2} will provide radial velocities for the
brightest stars ($G < 12$; \citealt{Gaia-Collaboration:2016}).

With large samples of candidate pairs, several groups have measured the
separation distribution of wide binaries at angular separations $\Delta \theta
\lesssim 1^\circ$ where the false-positive rate is expected to be lower
and the samples are likely predominantly wide binaries
(\citealt{Chaname:2004,Lepine:2007,Sesar:2008}).
In most cases, the angular or projected separation distribution is fit using a
decreasing power-law, and the power-law index, $\alpha$, is found to be $\alpha
\approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations.
Tentative reports of a steepening or cutoff of the separation distribution at
larger separations have been claimed (e.g., \citealt{Yoo:2004,Quinn:2009}), but
these claims are limited by small numbers of confirmed pairs at large
separations.

Most measurements of the separation distribution have neglected the possible
existence of larger separation ($\gtrsim 1~\pc$) unbound or ionized pairs.
Early theoretical work either explicitly ignored these stars (e.g.,
\citealt{Weinberg:1987}), or suggested that unbound stars would drift and no
longer remain spatially coherent within one Galactic orbit (e.g., see Appendix
in \citealt{Yoo:2004}).
More recent simulations of the disruption and evolution of wide binary
populations include an approximation of the Galactic tidal field and instead
predict that ionized binaries could appear as a second peak or component in the
separation distribution of comoving stars at much larger separations (10--
$1000~\pc$; \citealt{Jiang:2010}).
Though these simulations don't include perturbers like GMCs or black holes,
this at least indicates that a simple, monotonic separation distribution may
be too simple of a model for the separation distribution of comoving stars in
our Galaxy.

Indeed, observational work that followed appears to find tentative evidence for
the existence of a second peak in the separation distribution of
radial-velocity-confirmed comoving stars (\citealt{Shaya:2011}).
Using a ... todo: describe their paper

To verify the existence of this unbound population of very wide-separation
comoving pairs, we will need to identify a large sample of confirmed comoving
stars.
This will already be possible for the brightest stars with \gaia\ \DR{2}, but
obtaining even larger samples will require significant efforts to obtain radial
velocities for candidate comoving pairs.
In this work, we present a catalog of confirmed comoving star pairs discovered
through low-resolution spectroscopic follow-up of candidate comoving stars
selected from inferences using astrometric measurements from the \tgas\ catalog
alone (\citealt{Oh:2017}).

We describe the data, observations, and spectroscopic reduction in
\sectionname~\ref{sec:data}.
If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
and \sectionname~\ref{sec:discussion} where we present and discuss our main
results.
